Extended Data

### 2021-11-13 doc/manuscript/Dow_manuscript_Nature.Rmd
Update Dow_manuscript_Nature.Rmd
More edits

William McShea^1^

@@

Both MODIS-derived canopy foliage phenology and dendrometerband measurements of stem growth phenology generally shifted backwards as spring temperatures increased (Fig. `r DOY_timing`, `r EDF_pheno_all`).

We found a consistent effect of temperature ($T_{max}$ or $T_{min}$) throughout the Spring, but the strongest effects on stem-growth phenology were found using $T_{max}$ during a critical temperatrue window (CTW). CTW was identified by measuring the correlation between all combinations of weekly $T_{max}$ and $DOY_{25}$ from January 1 and $DOY_{25}$ (`r EDF_climwin`). The CTW was defined as the week(s) which had the strongest correlation with $DOY_{25}$.

@@

To understand how annual growth increments have responded to spring temperatures at the centennial scale, we analyzed tree-ring chronologies of `r n_chronologies_SCBI` species at SCBI[@helcoski_growing_2019] and `r n_chronologies_HF` species at Harvard Forest (`r EDT_species`), along with an additional `r n_chronologies-n_chronologies_SCBI-n_chronologies_HF` chronologies from `r n_cores_sites-2` sites (Fig. `r tree_rings`; `r EDF_map`; `r EDT_tree_ring_summary`)[@dorangeville_drought_2018].
In total, our analysis included `r n_chronologies` chronologies representing `r n_cores_species` broadleaf species at `r n_cores_sites` sites distributed from Alabama (Lat = 34.35) to Michigan (Lat = 45.56) and spanning a 15 $^\circ$C range in April $T_{max}$.
Across all chronologies, RWI was significantly (at p=0.05) positively correlated with April $T_{max}$ for only 1% of chronologies: 1 of 142 ring-porous and 2 of 66 diffuse-porous species-site combinations (`r EDT_tree_ring_summary`).
In contrast, RWI was frequently negatively correlated with $T_{max}$ during peak growing season months (May-August), with significant correlations for 53% (May: 47/142, Jun: 108/142, Jul: 90/142, Aug: 57/142) and 46% (May: 11/66, Jun: 53/66, Jul: 37/66, Aug: 22/66) of species-site-month combinations for ring-porous and diffuse-porous species, respectively.
$T_{min}$ generally exibited weaker relationships to annual growth than $T_{max}$, with few significant correlations between spring $T_{min}$ and RWI (`r EDF_tree_rings_TMIN`).

@@

# Acknowledgements {-#Acknowledgements} 
- people: (everyone who helped collect data-- see list on GitHub), Jess Shue (data)

- Funding: ForestGEO

**COAUTHORS, PLEASE ADD.**

@@

>>>>

William J. McShea^1^

@@

Both MODIS-derived canopy foliage phenology and dendrometerband measurements of stem growth phenology generally shifted backwards as spring temperatures increased (Fig. `r DOY_timing`, `r EDF_pheno_all`). We found a consistent effect of temperature ($T_{max}$ or $T_{min}$) throughout the Spring, but the strongest effects on stem-growth phenology were found using $T_{max}$ during a critical temperatrue window (CTW). CTW was identified by measuring the correlation between all combinations of weekly $T_{max}$ and $DOY_{25}$ from January 1 and $DOY_{25}$ (`r EDF_climwin`). The CTW was defined as the week(s) which had the strongest correlation with $DOY_{25}$.

@@

To understand how annual growth increments have responded to spring temperatures at the centennial scale, we analyzed tree-ring chronologies of `r n_chronologies_SCBI` species at SCBI[@helcoski_growing_2019] and `r n_chronologies_HF` species at Harvard Forest (`r EDT_species`), along with an additional `r n_chronologies-n_chronologies_SCBI-n_chronologies_HF` chronologies from `r n_cores_sites-2` sites (Fig. `r tree_rings`; `r EDF_map`; `r EDT_tree_ring_summary`)[@dorangeville_drought_2018].
In total, our analysis included `r n_chronologies` chronologies representing `r n_cores_species` broadleaf species at `r n_cores_sites` sites distributed from Alabama (Lat = 34.35) to Michigan (Lat = 45.56) and spanning a 15 $^\circ$C range in April $T_{max}$. We converted the ring-width measurements into a dimensionless growth index (hereafter simply referred to as RWI; see Methods).
Across all chronologies, RWI was significantly (at p=0.05) positively correlated with April $T_{max}$ for only 1% of chronologies: 1 of 142 ring-porous and 2 of 66 diffuse-porous species-site combinations (`r EDT_tree_ring_summary`).
In contrast, RWI was frequently negatively correlated with $T_{max}$ during peak growing season months (May-August), with significant correlations for 53% (May: 47/142, Jun: 108/142, Jul: 90/142, Aug: 57/142) and 46% (May: 11/66, Jun: 53/66, Jul: 37/66, Aug: 22/66) of species-site-month combinations for ring-porous and diffuse-porous species, respectively.
$T_{min}$ generally exibited weaker relationships to annual growth than $T_{max}$, with few significant correlations between spring $T_{min}$ and RWI (`r EDF_tree_rings_TMIN`).

@@

# Acknowledgements {-#Acknowledgements} 
- people: (everyone who helped collect data-- see list on GitHub), Jess Shue (data), We would like to thank Tsun Fung Au, Joshua Bregy, James Dickens, Karen Heeter, Anna Hennage, Daniel King, James McGee, Benjamin Lockwood, Josh Oliver, Karly Schmidt-Simard, Brandon Strange, Brynn Taylor, Michael Thornton, Senna Robeson, Matt Wenzel, and Luke Wylie
for assistance in the field and laboratory.

- Funding: ForestGEO, the collection of tree-ring samples was  funded by a USDA Agriculture and Food Research Initiative grant 2017-67013-26191 and from the Indiana University Vice Provost for Research Faculty Research Program.

-David Orwig
**COAUTHORS, PLEASE ADD.**

@@

*I can't figure out how to get this into zotero yet :(*
Stan Development Team. 2019. Stan Modeling Language Users Guide and Reference Manual, 2.28. https://mc-stan.org
